---
title: "PenguinsFlipperLengths ANOVA"
author: "Anna PerezWelter"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in .csv data. 
Packages: tidyverse, car, onewaytests
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
View(penguins)
```

Question: Is there as signif diff in mean flipper lengths for female Chinstrap, Gentoo and Adiele Penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA. 
-Indep. Observations
-Equal Variance (Levene's Test)
-Normality (histograms/qqplots)

```{r}
counts <- with(penguins, table(species, sex))
#View(counts)

penguins2 <- penguins %>%
  mutate(
    common_name = case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo", 
species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie", 
    )
  ) %>%
  select(-species) %>%
  filter(sex == "FEMALE")

View(penguins2)
```

Exploratory Graphs & Levene's Test for equal variances
```{r, message=FALSE}

hists <- ggplot(penguins2, aes(x=flipper_length))+
  geom_histogram(aes(fill=common_name))
hists

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name))+
  facet_wrap(~common_name, scale="free")
qqs

#Based on qqplots and histograms, observations for each group appear normally distributed. Also, n . 30 for all groups, so meand will be normally distr by Central Limit Theorem (CLT).

```

Evaluate Variances and perform Levene's Test for equal variances. 
```{r}
#If the largest sample cariances is < 4x frater than the smalest sample variance, then these are "close enough" to consider variances to be equal. 

variances <- penguins2 %>%
  group_by(common_name) %>%
  summarize(
    variance =var(flipper_length)
  )
variances # largest variance (33.1) definitely not more than 4x greater than the smallest variance (15.2)

#Levene's Test
#leveneTest function in 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
levene_penguin
#HO: No diff in variances (v are equal)
#HA: Variances are not equal. 

#There is no signif. differences in variances across species. 
```
One-Way ANOVA for penguin flipper lengths by species (common_name)

```{r}
#H0: Means across all groups are equal. 
#HA: At least two group means differ significantly

penguin_aov <- aov(flipper_length ~common_name, data=penguins2)
summary(penguin_aov)

#Post-hoc testing by Tukeyy's HSD

#H0: No signif diff in pairwise means
#HA: Signif diff in pairwise means
penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

#aov.test function in the 'onewaytests' function allows for inline referencing 
```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df)=statistic, p<0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD)